Generation of TypeScript Declaration Files from JavaScript Code

### 2020-01-03 ecoop-2020/main.tex
intro & motivation

This work presents \lstinline{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the NPM
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were generated and 
compared with the declaration file provided on DefinitelyTyped.
33 files out of 244 had no differences.
% but what about the remaining generated files
\end{abstract}

\section{Introduction}
JavaScript has become the most popular language for writing web applications \cite{github-statistics}. It is also gaining popularity for back-end applications running in NodeJS. Its dynamic typing speeds up programming enabling developers to create simple pieces of code very fast, making JavaScript a very appealing programming language.

JavaScript is being used for creating complex and large applications. However, JavaScript was not intended to be more than a scripting language. Maintaining and evolving large JavaScript codebases is notably challenging. Mistakes such as mistyped property names and misunderstood or unexpected type coercion cause developers to spend a significant amount of time in debugging sessions. A JavaScript code blog\footnote{https://wtfjs.com} compiles experiences from developers facing unexpected situations while programming in JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some of these unintuitive JavaScript behaviors.

\begin{lstlisting}[
    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion.},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float,
    abovecaptionskip=-\medskipamount
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
"01" < "00100"; // false
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
\end{lstlisting}

The overhead produced by such dynamic typing is not present in other languages that use build tools based on type information. The situation motivated the creation of TypeScript, a superset of JavaScript with typed annotations \cite{typescript}. It has become a widely used alternative among JavaScript developers, since it incorporates features that are helpful for developing and maintaining large applications \cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection of run-time errors detection and the integration of code intelligence tools like auto-completion in the IDEs.

Existing JavaScript libraries can be used in a TypeScript project by adding a declaration file that contains a typed description of the library's API. Declaration files are stored in a repository called DefinitelyTyped that contains declaration files for more than 6000 JavaScript libraries \cite{definitely-typed-repository}. Unfortunately, declaration files need to be manually created and maintained, which is error prone and time consuming. TypeScript does not perform any run-time check on these declaration files. A discrepancy between the declaration file and its corresponding JavaScript library would lead to additional frustration and debugging sessions, since type checks and code-intelligence features would be inaccurate.

Some previous work tackled the problem of automatically searching for mismatches between a declaration file and implementation code \cite{DBLP:conf/oopsla/FeldthausM14}. TSTest adapted the feedback-directed random testing technique, mainly used for testing Java libraries, to detect discrepancies between a declaration file and a JavaScript library \cite{DBLP:journals/pacmpl/KristensenM17}. Tools like TSInfer and TSEvolve are designed for assisting the creation of new declaration files and supporting the evolution the declaration file when the corresponding JavaScript library gets modified \cite{DBLP:conf/fase/KristensenM17}, respectively. TypeScript itself developed \lstinline[language={}]{dts-gen}, a tool that generates a declaration file that is meant to be used only as a `starting point for writing a high-quality declaration file' \cite{dts-gen}.

We explore in this work the possibilities for improving the existing tools. We present the tool \lstinline[language={}]{dts-generate}. It is based on an architecture that supports the generation of declaration files for an existing JavaScript library published to the NPM registry. The tool will gather data flow and type information at run-time and generate a declaration file based on that information.

The architecture supports the future incorporation of a Symbolic Execution Engine that expands the initial code base using the signatures in the declaration file. The iterative process of exploring new execution paths will refine the generated declaration file in each iteration.

Finally, we generated declaration files for 244 JavaScript libraries and evaluated them against DefinitelyTyped.

The contributions of this paper are as follows.
\begin{itemize}
    \item We introduce an architecture that supports generating TypeScript Declaration Files for a given JavaScript Library using run-time information. The architecture supports the future incorporation of a Symbolic Execution Engine that expands the initial code base using the signatures in the declaration file. The iterative process of exploring new execution paths will refine the generated declaration file in each iteration.

    \item We present the tool \lstinline{dts-generate}, a command line application that generates a valid TypeScript Declaration File for a specific NPM package using run-time information.

    \item Next, we deliver a test framework that supports generating a TypeScript Declaration File using \lstinline{dts-generate} for all JavaScript libraries in the DefinitelyTyped Repository. We present the results of running \lstinline{dts-generate} for 244 of these libraries.

    \item Finally, we present a TypeScript Declaration Files parser and comparator. Both tools were necessary for evaluating the results against DefinitelyTyped.
\end{itemize}

\section{Motivating Example}
The NPM package \lstinline{route-parser} is a simple route parsing, matching, and reversing library for Javascript\footnote{https://www.npmjs.com/package/route-parser}. It has about 35000 weekly downloads and 221 NPM packages depend on it. If a developer is creating or extending a JavaScript library written in TypeScript that depends on \lstinline{route-parser}, the TypeScript compiler and IDEs will need a declaration file for that JS library in order to perform static checking and code completion, respectively. We use \lstinline{dts-generate} to automatically generate a TypeScript Declaration File for this library. The tool will download the npm package, run it, gather run-time information and generate a valid TypeScript Declaration File. \figref{fig:motivating-example-route-parser-vscode} shows the generated declaration file for package \lstinline{route-parser} using \lstinline{dts-generate}. The result is a valid declaration file that can be used in a TypeScript project. For example, it can be seen in \figref{fig:motivating-example-route-parser-vscode} that Visual Studio Code performs code completion properly. Finally, if \lstinline{route-parser} gets modified in a future, a new declaration file could be automatically generated using \lstinline{dts-generate}.

@@

  \caption{\textbf{Declaration file for \lstinline{route-parser} generated with \lstinline{dts-generate}} - Constructor and methods are correctly identified. Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
  \label{fig:motivating-example-route-parser-vscode}
\end{figure}

\section{dts-generate - Generation of TypeScript Declaration Files}
We introduce \lstinline{dts-generate}, a command line tool which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}. The tool is intended to be used on existing, published \lstinline{npm} packages. The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.

@@

The command line interface was inspired in the \lstinline{dts-gen} \cite{dts-gen} package. It can be seen in \coderef{code:dts-generate-example} that invoking the package is very simple and the only required argument is the name of the module published to the npm registry.

\begin{lstlisting}[
    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \lstinline{abs}.},
    language=bash,
	label=code:dts-generate-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

@@

Obtaining code examples for a specific NPM Package is achieved in three steps:
\begin{itemize}
  \item Obtain the repository url from the package: The command \lstinline{npm view <PACKAGE> repository.url} can be used for retrieving the url of the package's repository.

@@

  \item Extract the code examples from the README file: Readme files are written using Markdown\footnote{https://www.markdownguide.org}, a very common lightweight and simple markup language. It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax. The Markdown identifiers for JavaScript are \lstinline{js} or \lstinline{javascript}. The code examples are finally retrieved by filtering the content within the corresponding code blocks.
\end{itemize}

@@

The implementation stores these observations as entities called \lstinline{interactions}. They are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation. The run-time information is finally returned as a JSON file that can be used for later processing. The tool is written in JavaScript and runs in Node.js within a Docker container.

@@

The tool is called \lstinline{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.

@@

\paragraph*{Repositories URL}
The url of the repositories could be retrieved for only 4974 modules. More than 1000 modules do not have the repository entry in their corresponding \lstinline{package.json} files. Therefore, the \lstinline{npm view <module> repository.url} command returns an empty value. This is even happening for important modules like \lstinline{ace}.

\paragraph*{Readme Files}
700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \lstinline{readme.md} or \lstinline{README.md}.

\paragraph*{Code Examples Extraction}
The 50\% loss is mainly explained because developers did not wrap their code around a block using the \lstinline{javascript} or \lstinline{js} tags. Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.

\paragraph*{Code Examples Execution}
The 2260 extracted code examples were executed by installing the required packages and running the code as a \lstinline{node} application. Working and functional code examples could only be extracted for 1448 modules. 812 modules did not run correctly and were discarded. Some failing samples were analyzed and there were mainly two reasons for the failure:
\begin{enumerate}
	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
\end{enumerate}

@@

\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \lstinline{abs}, \lstinline{dirname-regex} and \lstinline{escape-html}. All of them were generated using the \lstinline{module-function} template. There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.

@@

    \caption{\textbf{Module-function results} - Results are shown for modules \lstinline{abs}, \lstinline{dirname-regex}, \lstinline{escape-html}. On the left side the generated declaration file with \lstinline{dts-generate}. On the right side the corresponding file in the DefinitelyTyped repository. Functions are correctly detected and input types are accurately inferred. Both files are parsed for comparison, as explained in \secref{sec:dts-generate-evaluation}. Therefore, subtle differences in the syntax between both files are not important.}
\end{figure}

Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \lstinline{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

@@

    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \lstinline{ConstructorOptions} are correctly detected. Name of interface differs since it is automatically generated based on the name of the argument variable. Optional properties were not implemented, hence the \lstinline{undefined} type for some properties. Analogously, callback \lstinline{cb} is inferred as \lstinline{undefined}.}
    \label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

@@

It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct. For example, for \lstinline{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped. However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.

@@

    \caption{\textbf{Missing properties | datadog-metrics} - Properties \lstinline{aggregator} and \lstinline{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
    \label{fig:experiments-results-module-datadog-metrics}
\end{figure}

@@

\section{Related Work}
\paragraph*{Microsoft's dts-gen}
Microsoft developed \lstinline[language={}]{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}. Its documentation states that the result is however intended to be only used a starting point. The outcome needs to be refined afterwards by the developers.

@@

Formal aspects like applying the right template and using the correct syntax are perfectly covered by \lstinline[language={}]{dts-gen}.

@@

\section{Conclusion}
We have presented \lstinline{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library. The tool downloads code samples written by the developers from the library's repository. It uses these samples to execute the library and gather data flow and type information. The tool finally generates a TypeScript declaration file based on the information gathered at run-time.

>>>>

This work presents \texttt{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the NPM
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were generated and 
compared with the declaration file provided on DefinitelyTyped.
33 files out of 244 had no differences.
% but what about the remaining generated files
\end{abstract}

\section{Introduction}
\label{sec:introduction}
JavaScript has become the most popular language for writing web
applications \cite{github-statistics}. It is also gaining popularity
for back-end applications running in NodeJS, a JavaScript-based
server-side platform. JavaScript is appealing to developers because
its forgiving dynamic typing enables 
them to create simple pieces of code very quickly and proceed on a
trial-and-error basis.

JavaScript was never intended to be more than a
scripting language. Hence, it lacks features for maintaining and evolving large
codebases.
Presently, however, JavaScript is being used for creating large and complex
applications. 
Mistakes such as mistyped property
names and misunderstood or unexpected type coercions cause developers
to spend a significant amount of time in debugging. There is ample
evidence for such mishaps. For example, a 
JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
from developers facing unexpected situations while programming in
JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
of these unintuitive JavaScript behaviors.

\begin{lstlisting}[
    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float,
    abovecaptionskip=-\medskipamount
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
\end{lstlisting}
% pjt: why unexpected?
% "01" < "00100"; // false

The cognitive load produced by such dynamic typing (which includes
unchecked property names) is not present in other languages that use
build tools based on type information. This insight motivated the
creation of TypeScript, a superset of JavaScript with type
annotations \cite{typescript}. It has become a widely used alternative
among JavaScript developers, because it incorporates features that are
helpful for developing and maintaining large applications
\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
of several kinds of run-time errors and the integration of code intelligence
tools like auto-completion in an IDE.

Existing JavaScript libraries can be used in a TypeScript project by
adding a declaration file that contains a description of the library's
API in terms of types. Lots of declaration files for popular
JavaScript libraries have been created in a community effort in the
DefinitelyTyped repository~\cite{definitely-typed-repository}.
At the time of writing this repository contains declaration files for
more than 6000 JavaScript libraries. Unfortunately, most declaration
files in this libraries have been manually created and maintained,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the inaccurate code-intelligence features are misleading the
programmer. Moreover, TypeScript does not perform any run-time 
checking on types, either, so that a discrepancy  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such behavior can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

Some previous work tackled the problem of improving the quality of
declaration files. Feldthaus et al
\cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
mismatches between a declaration file and implementation code. TSTest
\cite{DBLP:journals/pacmpl/KristensenM17}
adapted feedback-directed random testing to detect discrepancies between a
declaration file and a JavaScript library. Tools like TSInfer and
TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
assisting the creation of new declaration 
files and supporting the evolution the declaration file when the
corresponding JavaScript library gets modified, respectively. They
rely on an existing static analyzer for JavaScript \cite{TAJS}.
TypeScript itself
developed \texttt{dts-gen}, a tool that generates a
declaration file that is meant to be used as a \emph{starting point
for writing a high-quality declaration file} \cite{dts-gen}.

In this work we present the tool \texttt{dts-generate}
as a first step to explore the possibilities for
generating useful declaration files without the heavy lifting of
static analysis. \texttt{dts-generate} comes with a framework that
supports the generation of declaration files for an existing
JavaScript library published to the NPM registry. The tool gathers
data flow and type information at run-time to generate a declaration
file based on that information.

The novelty of our tool is twofold:
\begin{enumerate}
\item 
  We do not rely on static analysis, which is hard to implement
  soundly and precisely and which is prone to maintenance problems
  when keeping up with JavaScript's yearly language updates.
\item
  Instead we extract example code from the programmer's library
  documentation and rely on dynamic analysis to extract typed usage
  patterns for the library from the example runs.
\end{enumerate}
% The architecture supports the future incorporation of a Symbolic
% Execution Engine that expands the initial code base using the
% signatures in the declaration file. The iterative process of exploring
% new execution paths will refine the generated declaration file in each
% iteration.

% Finally, we generated declaration files for 244 JavaScript libraries
% and evaluated them against the DefinitelyTyped repository.

The contributions of this paper are as follows.
\begin{itemize}
% \item We introduce an architecture that supports generating TypeScript
%   Declaration Files for a given JavaScript Library using run-time
%   information. The architecture supports the future incorporation of a
%   Symbolic Execution Engine that expands the initial code base using
%   the signatures in the declaration file. The iterative process of
%   exploring new execution paths will refine the generated declaration
%   file in each iteration.
\item A framework that extracts code examples from the
  documentation of an NPM modules and collects run-time type
  information from running these examples.

\item The tool \texttt{dts-generate}, a command line
  application that generates a valid TypeScript declaration file for a
  specific NPM package using run-time information.

\item A comparator for TypeScript declaration files. This tool is
  necessary for evaluating our framework and also useful to detect
  incompatibilities when evolving JavaScript modules.
\item An evaluation of our framework. We examined all 6029 entries in
  the DefinitelyTyped repository and found 244 sufficiently
  well-documented NPM libraries, on which we ran \texttt{dts-generate}
  and compared the outcome with the respective declaration file from the
  DefinitelyTyped repository.

\end{itemize}

\section{Motivating Example}
The NPM package \texttt{route-parser} is a simple route parsing,
matching, and reversing library for
Javascript\footnote{https://www.npmjs.com/package/route-parser}. It
has about 35000 weekly downloads and 221 NPM packages depend on it. If
a developer creates or extends TypeScript code that depends on the
\texttt{route-parser} library, the TypeScript
compiler and IDE requires a declaration file for that library to
perform static checking and code completion, respectively. We 
use \texttt{dts-generate} to automatically generate a TypeScript
declaration file for \texttt{route-parser}. The tool downloads the NPM 
package, runs the examples extracted from its documentation, gathers
run-time information, and generates a TypeScript declaration
file. The result is a valid declaration file
shown in \figref{fig:motivating-example-route-parser-vscode}  that 
can be used in a TypeScript project. For example,  Visual Studio
Code performs code completion properly
(\figref{fig:motivating-example-route-parser-vscode}). If the
\texttt{route-parser} package gets modified in a future, a new declaration
file can be generated automatically using
\texttt{dts-generate}. Our comparator tool can compare the new file
for incompatibilities with the previous declaration file.

@@

  \caption{\textbf{Declaration file for \texttt{route-parser} generated with \texttt{dts-generate}} - Constructor and methods are correctly identified. Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
  \label{fig:motivating-example-route-parser-vscode}
\end{figure}

\section{dts-generate - Generation of TypeScript Declaration Files}
We introduce \texttt{dts-generate}, a command line tool which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}. The tool is intended to be used on existing, published \texttt{npm} packages. The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.

@@

The command line interface was inspired in the \texttt{dts-gen} \cite{dts-gen} package. It can be seen in \coderef{code:dts-generate-example} that invoking the package is very simple and the only required argument is the name of the module published to the npm registry.

\begin{lstlisting}[
    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \texttt{abs}.},
    language=bash,
	label=code:dts-generate-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

@@

Obtaining code examples for a specific NPM Package is achieved in three steps:
\begin{itemize}
  \item Obtain the repository url from the package: The command \texttt{npm view <PACKAGE> repository.url} can be used for retrieving the url of the package's repository.

@@

  \item Extract the code examples from the README file: Readme files are written using Markdown\footnote{https://www.markdownguide.org}, a very common lightweight and simple markup language. It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax. The Markdown identifiers for JavaScript are \texttt{js} or \texttt{javascript}. The code examples are finally retrieved by filtering the content within the corresponding code blocks.
\end{itemize}

@@

The implementation stores these observations as entities called \texttt{interactions}. They are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation. The run-time information is finally returned as a JSON file that can be used for later processing. The tool is written in JavaScript and runs in Node.js within a Docker container.

@@

The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.

@@

\paragraph*{Repositories URL}
The url of the repositories could be retrieved for only 4974 modules. More than 1000 modules do not have the repository entry in their corresponding \texttt{package.json} files. Therefore, the \texttt{npm view <module> repository.url} command returns an empty value. This is even happening for important modules like \texttt{ace}.

\paragraph*{Readme Files}
700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \texttt{readme.md} or \texttt{README.md}.

\paragraph*{Code Examples Extraction}
The 50\% loss is mainly explained because developers did not wrap their code around a block using the \texttt{javascript} or \texttt{js} tags. Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.

\paragraph*{Code Examples Execution}
The 2260 extracted code examples were executed by installing the required packages and running the code as a \texttt{node} application. Working and functional code examples could only be extracted for 1448 modules. 812 modules did not run correctly and were discarded. Some failing samples were analyzed and there were mainly two reasons for the failure:
\begin{enumerate}
	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
\end{enumerate}

@@

\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \texttt{abs}, \texttt{dirname-regex} and \texttt{escape-html}. All of them were generated using the \texttt{module-function} template. There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.

@@

    \caption{\textbf{Module-function results} - Results are shown for modules \texttt{abs}, \texttt{dirname-regex}, \texttt{escape-html}. On the left side the generated declaration file with \lstinline{dts-generate}. On the right side the corresponding file in the DefinitelyTyped repository. Functions are correctly detected and input types are accurately inferred. Both files are parsed for comparison, as explained in \secref{sec:dts-generate-evaluation}. Therefore, subtle differences in the syntax between both files are not important.}
\end{figure}

Templates of type \lstinline{module-class} are shown for modules \texttt{flake-idgen}, \texttt{route-parser} and \texttt{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \texttt{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

@@

    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \texttt{ConstructorOptions} are correctly detected. Name of interface differs since it is automatically generated based on the name of the argument variable. Optional properties were not implemented, hence the \lstinline{undefined} type for some properties. Analogously, callback \texttt{cb} is inferred as \lstinline{undefined}.}
    \label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

@@

It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct. For example, for \texttt{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped. However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.

@@

    \caption{\textbf{Missing properties | datadog-metrics} - Properties \texttt{aggregator} and \texttt{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
    \label{fig:experiments-results-module-datadog-metrics}
\end{figure}

@@

\section{Related Work}
\paragraph*{Microsoft's dts-gen}
Microsoft developed \texttt{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}. Its documentation states that the result is however intended to be only used a starting point. The outcome needs to be refined afterwards by the developers.

@@

Formal aspects like applying the right template and using the correct syntax are perfectly covered by \texttt{dts-gen}.

@@

\section{Conclusion}
We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library. The tool downloads code samples written by the developers from the library's repository. It uses these samples to execute the library and gather data flow and type information. The tool finally generates a TypeScript declaration file based on the information gathered at run-time.
